---
title: "RBMRB"
author: "Kumaran Baskaran"
date: "November 08, 2016"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

#RBMRB
BMRB collects, annotates, archives, and disseminates (worldwide in the public domain) the important spectral and quantitative data derived from NMR spectroscopic investigations of biological macromolecules and metabolites. The goal is to empower scientists in their analysis of the structure, dynamics, and chemistry of biological systems and to support further development of the field of biomolecular NMR spectroscopy

RBMRB is a library to fetch NMR chemical shift data directly from [BMRB](http://bmrb.wisc.edu/) into R environment 
as a data frame in R. This facilitates access to BMRB data for statistical analysis and data visualization. It is using the [BMRB-API](https://github.com/uwbmrb/BMRB-API)  to fetch the data from BMRB database. 

##Installation
RBMRB library has been developed and tested in  R version 3.3.x.  It requires the following R packages preinstalled


 + **httr** to import data from BMRB web server(version 1.2.1 or later)
 + **data.table** to format the imported data into a data frame in R (version 1.9.6 or later)
 + **rjson** to deal with BMRB-API (version 0.2.15 or later)
 + **ggplot2** to simulate spectra (version 2.1.0 or later)
 + **plotly** for interactive graphics in simulated spectra (version 4.5.2 or later)
 
Users should make sure that the above packages have beed installed correctly with the required versions, before proceeding to RMBRM insallation.

Here is the instruction to install those packages. Open your R and use the following command 
```{r eval=FALSE}
installed.packages(c("httr","data.table","rjson","ggplot2","plotly"))
```

Once the necessary packages have been installed, proceed with RMRBM installation. The source file can be downloaded from [GitHub](https://github.com/uwbmrb/RBMRB/raw/master/RBMRB_1.0.tar.gz)

After downloading the source file, use the following command to install RBMRB library
```{r eval=F}
install.packages("~/Downloads/RBMRB_1.0.tar.gz",repos=NULL,type="source")
```

Note: provide the correct path to the downloaded file.

##Usage
RBMRB can be used in a similar way like any other library in R. 

```{r}
library(RBMRB)
```

##Data access
BMRB data can be imported in two ways

+ **Entry method** Chemical shift data from single or multiple entries
+ **Atom method** Chemical shift data from all entries for a given atom

###Entry method
####fetch_entry_chemical_shifts:
This function will fetch the 'Atom_chem_shift' loop from a NMR-STAR file for a given entry or a list of entries in CSV format. This function works on both macromolecules and metabolites data base. For metabilites entry ids should have right prefix (example 'bmse000034')

######Examples:
```{r}
df1<-fetch_entry_chemical_shifts(15060)
df2<-fetch_entry_chemical_shifts(c(17074,17076,17077))
df2<-fetch_entry_chemical_shifts(c('17074','17076','17077'))
df3<-fetch_entry_chemical_shifts(c('bmse000034','bmse000035','bmse000036'))
```

These data frames have the following columns
```{r}
colnames(df1)
```

###Atom method
####fetch_atom_chemical_shifts:
This function will fetch the chemical shift data from all the entries for a given atom. The atom name should be in NMR-STAR atom nomenclature.

######Examples:
```{r}
df4<-fetch_atom_chemical_shifts('CG2')
df5<-fetch_atom_chemical_shifts('C9')
```

These data frames have the following columns
```{r}
colnames(df4)
```

##Data manipulation
There are few data manipulation functions are availbale to facilitate plotting. 
####convert_cs_to_n15hsqc:
This function will reformat the chemical shift data frame into a data frame which is easy to plot the N15-HSQC spectrum from the data. 

####Examples
```{r}
n15hsqc1<-convert_cs_to_n15hsqc(df1)
n15hsqc2<-convert_cs_to_n15hsqc(df2)
library(ggplot2)
plt1<-ggplot(n15hsqc1)+geom_point(aes(x=H,y=N))
plt1
plt2<-ggplot(n15hsqc2)+geom_point(aes(x=H,y=N,color=Entry_ID))
plt2
plt3<-ggplot(n15hsqc2)+geom_point(aes(x=H,y=N,color=Entry_ID))+geom_line(aes(x=H,y=N,group=Comp_index_ID))
plt3
```

####convert_cs_to_c13hsqc:
This function will reformat the chemical shift data frame into a data frame which is easy to plot the C13-HSQC spectrum from the data. 

####Examples
```{r}
c13hsqc1<-convert_cs_to_c13hsqc(df1)
c13hsqc2<-convert_cs_to_c13hsqc(df2)
library(ggplot2)
plt1<-ggplot(c13hsqc1)+geom_point(aes(x=H,y=C))
plt1
plt2<-ggplot(c13hsqc2)+geom_point(aes(x=H,y=C,color=Entry_ID))
plt2
```

####convert_cs_to_tocsy:
This function will reformat the chemical shift data frame into a data frame which is easy to plot the TOCSY spectrum from the data. Note : Since both dimensions have protein chemical shifts, the columns are named as Val.x and Val.y

####Examples
```{r}
tocsy1<-convert_cs_to_tocsy(df1)
tocsy2<-convert_cs_to_tocsy(df2)
library(ggplot2)
plt1<-ggplot(tocsy1)+geom_point(aes(x=Val.x,y=Val.y))
plt1
plt2<-ggplot(tocsy2)+geom_point(aes(x=Val.x,y=Val.y,color=Entry_ID))
plt2
```

####filter_residue:
This function will filter the data frame and remove all non standard amino acids. The data frame should contain the amino acid information in the Comp_ID column. 
####Examples
```{r}
df6<-fetch_atom_chemical_shifts('CG2')
df7<-filter_residue(df6)
```

##Data visualization

RBMRB library contains few functions to generate interactive visualization of BMRB data with out any data manipulation. The interactive visualizations use **plotly** library. If user has problem with plotly, then he may disable it by providing an argument 'interactive=FALSE' for these functions. 

####HSQC_15N
This function will simulae N15-HSQC spectrum for a given entry or list of entries. 

#####Examples

These interactive visualization can be exported as single stand alone html file
```{r}
spec1<-HSQC_15N(15060)
spec1
spec2<-HSQC_15N(c(17074,17076,17077),type='line')
spec2
spec3<-HSQC_15N(c(17074,17076,17077),type='line',interactive = F)
spec3
```

####HSQC_13C
This function will simulae C13-HSQC spectrum for a given entry or list of entries. 

#####Examples

These interactive visualization can be exported as single stand alone html file
```{r}
spec1<-HSQC_13C(15060)
spec1
spec2<-HSQC_13C(c(17074,17076,17077))
spec2
spec3<-HSQC_13C(c(17074,17076,17077),interactive = F)
spec3
```

####TOCSY
This function will simulae TOCSY spectrum for a given entry or list of entries. 

#####Examples

These interactive visualization can be exported as single stand alone html file
```{r}
spec1<-TOCSY(15060)
spec1
spec2<-TOCSY(c(17074,17076,17077))
spec2
spec3<-TOCSY(c(17074,17076,17077),interactive = F)
spec3
```
