---
title: 'RBMRB : A package to download and visualize BMRB data in R'
author: "Kumaran Baskaran"
date: "November 08, 2016"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

#RBMRB
[BMRB](http://www.bmrb.wisc.edu/) collects, annotates, archives, and disseminates (worldwide in the public domain) the important spectral and quantitative data derived from NMR spectroscopic investigations of biological macromolecules and metabolites. The goal is to empower scientists in their analysis of the structure, dynamics, and chemistry of biological systems and to support further development of the field of biomolecular NMR spectroscopy

[RBMRB](https://github.com/uwbmrb/RBMRB) is a library to fetch NMR chemical shift data directly from [BMRB](http://www.bmrb.wisc.edu/) into R environment 
as a data frame in R. This facilitates access to BMRB data for statistical analysis and data visualization. It is using the [BMRB-API](https://github.com/uwbmrb/BMRB-API)  to fetch the data from BMRB database. 

##Installation
RBMRB library has been developed and tested in  R version 3.3.x.  It requires the following R packages preinstalled


 + **httr** to import data from BMRB web server(version 1.2.1 or later)
 + **data.table** to format the imported data into a data frame in R (version 1.9.6 or later)
 + **rjson** to deal with BMRB-API (version 0.2.15 or later)
 + **ggplot2** to simulate spectra (version 2.1.0 or later)
 + **plotly** for interactive graphics in simulated spectra (version 4.5.2 or later)
 
Users should make sure that the above packages have been installed correctly with the required versions, before proceeding to RBMRB installation.

Here is the instruction to install those packages. Open your R and use the following command 
```{r eval=FALSE}
installed.packages(c("httr","data.table","rjson","ggplot2","plotly"))
```

####Method 1
Once the necessary packages have been installed, proceed with RMRBM installation. The source file can be downloaded from [GitHub](https://github.com/uwbmrb/RBMRB/raw/master/RBMRB_2.0.tar.gz)

After downloading the source file, use the following command to install RBMRB library
```{r eval=F}
install.packages("~/Downloads/RBMRB_2.0.tar.gz",repos=NULL,type="source")
```

Note: provide the correct path to the downloaded file.

####Method 2
If you have devtools library in your R, then you can install directly from GitHub.
```{r eval=F}
library(devtools)
install_github("uwbmrb/RBMRB/RBMRB")
```
##Usage
RBMRB can be used in a similar way like any other library in R. 

```{r}
library(RBMRB)
```

##Data access
BMRB data can be imported in two ways

+ **Entry method** Chemical shift data from single or multiple entries
+ **Atom method** Chemical shift data from all entries for a given atom

###Entry method
####fetch_entry_chemical_shifts:
This function will fetch the 'Atom_chem_shift' loop from a NMR-STAR file for a given entry or a list of entries in CSV format. This function works on both macromolecules and metabolites data base. For metabolites entry ids should have right prefix (example 'bmse000034')

######Examples:
```{r}
df1<-fetch_entry_chemical_shifts(15060)
df2<-fetch_entry_chemical_shifts(c(17074,17076,17077))
df2<-fetch_entry_chemical_shifts(c('17074','17076','17077'))
df3<-fetch_entry_chemical_shifts(c('bmse000034','bmse000035','bmse000036'))
```

These data frames have the following columns
```{r}
colnames(df1)
```
Sample data output
```{r}
head(df1)
```
###Atom method
####fetch_atom_chemical_shifts:
This function will fetch the chemical shift data from all the entries for a given atom. The atom name should be in NMR-STAR atom nomenclature.

######Examples:
```{r}
df4<-fetch_atom_chemical_shifts('CG2')
df5<-fetch_atom_chemical_shifts('C9')
```

These data frames have the following columns
```{r}
colnames(df4)
```
Sample data output
```{r}
head(df4)
```

##Data manipulation
There are few data manipulation functions are available to facilitate plotting. 

####convert_cs_to_n15hsqc:
This function will reformat the chemical shift data frame into a data frame which is easy to plot the N15-HSQC spectrum from the data. 

####Examples
```{r}
n15hsqc1<-convert_cs_to_n15hsqc(df1)
n15hsqc2<-convert_cs_to_n15hsqc(df2)
```
The output data frame will look like
```{r}
head(n15hsqc1)
```

This data frame is easy to plot using any plotting library
```{r}
library(ggplot2)
plt1<-ggplot(n15hsqc1)+geom_point(aes(x=H,y=N))
plt1
plt2<-ggplot(n15hsqc2)+geom_point(aes(x=H,y=N,color=Entry_ID))
plt2
plt3<-ggplot(n15hsqc2)+geom_point(aes(x=H,y=N,color=Entry_ID))+geom_line(aes(x=H,y=N,group=Comp_index_ID))
plt3
```

####convert_cs_to_c13hsqc:
This function will reformat the chemical shift data frame into a data frame which is easy to plot the C13-HSQC spectrum from the data. 

####Examples
```{r}
c13hsqc1<-convert_cs_to_c13hsqc(df1)
c13hsqc2<-convert_cs_to_c13hsqc(df2)
```

The output data frame will look like
```{r}
head(c13hsqc1)
```
and the user may generate a spectrum using the following script
```{r}
library(ggplot2)
plt1<-ggplot(c13hsqc1)+geom_point(aes(x=H,y=C))
plt1
plt2<-ggplot(c13hsqc2)+geom_point(aes(x=H,y=C,color=Entry_ID))
plt2
```

####convert_cs_to_tocsy:
This function will reformat the chemical shift data frame into a data frame which is easy to plot the TOCSY spectrum from the data. Note : Since both dimensions have proton chemical shifts, the columns are named as Val.x and Val.y

####Examples
```{r}
tocsy1<-convert_cs_to_tocsy(df1)
tocsy2<-convert_cs_to_tocsy(df2)
```
after conversion the data will look like
```{r}
head(tocsy1)
```
Plotting TOCSY spectrum
```{r}
library(ggplot2)
plt1<-ggplot(tocsy1)+geom_point(aes(x=Val.x,y=Val.y))
plt1
plt2<-ggplot(tocsy2)+geom_point(aes(x=Val.x,y=Val.y,color=Entry_ID))
plt2
```

####filter_residue:
This function will filter the data frame and remove all non standard amino acids. The data frame should contain the amino acid information in the Comp_ID column. 
####Examples
```{r}
df6<-fetch_atom_chemical_shifts('CG2')
df7<-filter_residue(df6)
```

##Data visualization

RBMRB library contains few functions to generate interactive visualization of BMRB data with out any data manipulation. The interactive visualizations use **plotly** library. If user has problem with plotly, then this feature may be disabled by providing an argument 'interactive=FALSE' for these functions. These interactive plots can be zoomed in and out using a mouse and will show tooltip information when you mouse over. These visualizations can be exported as a stand alone html file

####HSQC_15N
This function will simulate N15-HSQC spectrum for a given entry or list of entries. 

#####Examples

These interactive visualization can be exported as single stand alone html file
```{r}
spec1<-HSQC_15N(15060)
spec1
spec2<-HSQC_15N(c(17074,17076,17077),type='line')
spec2
```
```{r}
spec3<-HSQC_15N(c(17074,17076,17077),type='line',interactive = F)
spec3
```

####HSQC_13C
This function will simulate C13-HSQC spectrum for a given entry or list of entries. 

#####Examples

These interactive visualization can be exported as single stand alone html file
```{r}
spec1<-HSQC_13C(15060)
spec1
spec2<-HSQC_13C(c(17074,17076,17077))
spec2
```
Non interactive plot
```{r}
spec3<-HSQC_13C(c(17074,17076,17077),interactive = F)
spec3
```

####TOCSY
This function will simulate TOCSY spectrum for a given entry or list of entries. 

#####Examples

These interactive visualization can be exported as single stand alone html file
```{r}
spec1<-TOCSY(15060)
spec1
spec2<-TOCSY(c(17074,17076,17077))
spec2
```
Non interactive plot
```{r}
spec3<-TOCSY(c(17074,17076,17077),interactive = F)
spec3
```

####chemical_shift_corr
This function will plot the distribution of chemical shift correlation between any two atoms from the 20 standard amino acids. The distribution of a particular residue may turn on and off by clicking the residue name in the legend.

```{r}
corr_plot1<-chemical_shift_corr('CB','N')
corr_plot1
corr_plot2<-chemical_shift_corr('CA','HA*')
corr_plot2
```
Non interactive plot
```{r}
corr_plot1<-chemical_shift_corr('CB','N',interactive = F)
corr_plot1
corr_plot2<-chemical_shift_corr('CA','HA*',interactive = F)
corr_plot2
```

##Bug Report
Please report the bugs to the [RBMRB GitHub repository](https://github.com/uwbmrb/RBMRB). Any new features can also be requested through the GitHub repository. 
